---
title: "Untitled"
output: html_document
date: "2026-01-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(readxl)
library(dplyr)
library(stringr)
library(writexl)
library(janitor)
```

# Load files
NeED TO CHECK:
## Progress: Charity tab should be included
## What happens with the early files - they are not captured in teh new column (not in the Progress probably)


# WORKFLOW
```{r}
# Goal: Harmonize and compare extraction workflow status between:
#   (a) the SUMMARY file (Extraction overview) and
#   (b) the Progress tracker workbook (multiple sheets),
# keyed on unique_study_id.
#
# 1) Read inputs
# a) Read the SUMMARY file (sheet 1) into `summary_status`.
# b) Read the Progress tracker across sheets 1–3 (no column names),
#      clean each sheet to a standard tabular format, and row-bind into
#      a single `progress` data frame.
#
# 2) Clean each Progress tracker sheet (inside lapply)
#   a) Remove known non-data rows (rows 1, 2, and 4).
#   b) Detect the header row by locating "unique_study_id" in column 1
#      (case- and whitespace-insensitive); stop with an informative error
#      if it is not found.
#   c) Apply the detected header row as column names, and keep only rows
#      below the header (the actual data region).
#   2) Remove non-study rows where column 1 contains "AsReview".
#
# 3) Build a status lookup from Progress tracker
#   a) Select and standardize key fields: `unique_study_id` and
#      `workflow_extract` (renamed to `status_new`), trimming whitespace.
#   b) Convert blank statuses to NA.
#   c) Deduplicate to one row per `unique_study_id` by keeping the first
#      non-missing `status_new` value (if any).
#
# 4) Merge into SUMMARY
#   a) Standardize `unique_study_id` in `summary_status`.
#   b) Left-join the Progress-based lookup to add `status_new` to
#      `summary_status` (preserving all summary records).
#   c) Relocate `status_new` next to the existing `status` column to
#      facilitate QA of status changes (status vs status_new).

```

#1-2
```{r}
# Summary file
summary_status <- read_excel("C:/Users/ANowak/OneDrive - CGIAR/PhD/Ch2/5.Extraction/Extraction overview-use this.xlsx", sheet = 1)

# Progress tracker
progress <- lapply(
  1:4,
  function(s) {

    df <- read_excel(
      "C:/Users/ANowak/OneDrive - CGIAR/AE-SysRev-shared/Progress tracker.xlsm",
      sheet = s,
      col_names = FALSE
    )

    # remove rows 1, 2, 4
    df <- df[-c(1, 2, 4), , drop = FALSE]

    # find header row (unique_study_id in column 1)
    col1 <- tolower(trimws(as.character(df[[1]])))
    header_row <- which(col1 == "unique_study_id")[1]

    if (is.na(header_row)) {
      stop(sprintf(
        "Header row with 'unique_study_id' not found on sheet %s.", s
      ))
    }

    # apply header
    header <- as.character(unlist(df[header_row, ], use.names = FALSE))
    df2 <- df[(header_row + 1):nrow(df), , drop = FALSE]
    names(df2) <- header

    # remove rows where column 1 contains 'AsReview'
    df2 <- df2[!grepl("AsReview", as.character(df2[[1]]), ignore.case = TRUE), ]

    df2
  }
) |> bind_rows()
```

# 3, 4


```{r}
# Build a lookup from PROGRESS: unique_study_id -> workflow_extract
progress_lookup <- progress |>
  transmute(
    unique_study_id = str_trim(as.character(unique_study_id)),
    status_new = str_trim(as.character(extract_status))
  ) |>
  mutate(status_new = na_if(status_new, "")) |>
  group_by(unique_study_id) |>
  summarise(
    status_new = dplyr::first(na.omit(status_new), default = NA_character_),
    .groups = "drop"
  )

# Join onto SUMMARY to create status_new in summary_status
summary_status <- summary_status |>
  mutate(unique_study_id = str_trim(as.character(unique_study_id))) |>
  left_join(progress_lookup, by = "unique_study_id") |>
  mutate(
    status = str_trim(as.character(status)),
    status = na_if(status, ""),
    status_new = coalesce(status_new, status)   # <-- fill only NAs in status_new
  ) |>
  relocate(status_new, .after = status)

```

# check for inconsistencies and misisng information
```{r}
## study ids in summary sheet and not in progress tracker
in_summary_only <- summary_status %>%
  anti_join(progress, by = "unique_study_id")
  
## study ids in progress tracker and not in summary
in_progress_only <- progress %>%
  anti_join(summary_status, by = "unique_study_id")

## records were status labels differ across status and status_new columns in the summary df
status_mismatch <- summary_status %>%
  filter(status != status_new)
```



# Status of extractions
```{r}
# summary of extractions
status <- summary_status %>%
  count(status_new, name = "n") %>%
  mutate(pct = 100 * n / sum(n))

# n experimental and quasi-experimental studies by status type
status_priority <- summary_status %>%
  mutate(
    status_grp = if_else(is.na(status_new), "NA", as.character(status_new)),
    has_label_new = !is.na(label_new) & as.character(label_new) != ""
  ) %>%
  filter(status_grp %in% c("completed","in progress","queued","stopped","NA")) %>%
  group_by(status_grp) %>%
  summarise(
    n_total = n(),
    n_with_label_new = sum(has_label_new),
    pct_with_label_new = 100 * n_with_label_new / n_total,
    .groups = "drop"
  )

#check where there are inconsistencies between the progress tracker and the overview
# Rows where status and status_new are inconsistent
mismatches <- summary_status %>%
  mutate(
    status     = str_trim(str_to_lower(as.character(status))),
    status_new = str_trim(str_to_lower(as.character(status_new))),

    # map status to the "status_new" vocabulary for comparison
    status_norm = case_when(
      status == "extracted" ~ "completed",   # treat as equivalent
      TRUE ~ status
    )
  ) %>%
  filter(!(is.na(status_norm) & is.na(status_new))) %>%   # both NA => match
  filter(status_norm != status_new)                       # mismatches only

mismatches

```


```{r}
# Save files
write.csv(summary_status, "C:/Users/ANowak/OneDrive - CGIAR/AE-SysRev-shared/WorkOverview/Summary-overview-tracker.csv")

#write.csv(in_progress_only, "C:/Users/ANowak/OneDrive - CGIAR/AE-SysRev-shared/WorkOverview/Papers-in-tracker-only.csv")

#write.csv(status_mismatch, "C:/Users/ANowak/OneDrive - CGIAR/AE-SysRev-shared/WorkOverview/Papers-in-overview-only.csv")

#write.csv(status_mismatch, "C:/Users/ANowak/OneDrive - CGIAR/AE-SysRev-shared/WorkOverview/Status-mismatch-overview-tracker.csv")
```





Worflow 
- Load "progress" file and only take into account sheets 1-3. This is the source file. This is an xlsm format. 
- From "progress" file, remove rows 1, 2, 4 from each sheet (these are part of the header table)
- Load "summary" file and only take into account sheet 1. This is the overview file
- In the "progress" df, create a column that says "status_new" 
based on unique_study_id column, which is common to both dataframes, I need to assign to the "status_new" column the text that appears in the respective cell in the "summary" file. for the summary file, need to take into account all 3 sheets. 