---
title: "Untitled"
output: html_document
date: "2026-06-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Check if all extracted/ QC papers marked as such in the Progress tracker are in the respective subfolders under Boris, Jabesh, Judy, Charity; 
# Identify which paper is misplaced 

Action: 
# Communicate findngs to the team
# place paper correctly (manually)

```{r}

library(readxl)

# ============================================================
# v1002 Misplaced File Check using Extract + QC sheets
#
# Purpose:
#   Identify papers that are in the wrong workflow folder.
#
# Important rule:
#   Folder matching is done by PAPER ID, not by full file version.
#   Example:
#      Tracker expects: v1002-JM00505.xlsm
#      Folder may contain: v0611-JM00505.xlsm
#   These are treated as the same paper because both are JM00505.
#
# Folder rules:
#   1. Not yet in QC sheet                           -> 2.Extracted
#   2. QC status = rejected                          -> 5.Rejected
#   3. QC status = queued                            -> 4.Queued
#   4. QC status = completed + final check yes-all   -> 3.Quality control
#   5. QC status = completed + zero/no comments      -> 3.Quality control
#   6. QC started but comments not resolved          -> 2.Extracted
#
# Also flags possible self-QC cases.
#
# Output:
#   v1002_misplaced_files_qc_logic.csv
# ============================================================

progress_file <- "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Progress tracker.xlsm"

extract_sheets_requested <- c("BS-Extract", "JM-Extract", "JO-Extract", "CJ-Extract")
qc_sheets_requested <- c("BS-QC", "JM-QC", "JO-QC", "CJ-QC")

base_paths <- c(
  BS = "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Boris",
  JO = "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Jabesh",
  JM = "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Judy",
  CJ = "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Charity"
)

output_file <- "v1002_misplaced_files_qc_logic.csv"

# -----------------------------
# Helper functions
# -----------------------------

clean_text <- function(x) {
  x <- as.character(x)
  x[is.na(x)] <- ""
  x <- gsub("[\r\n]+", " ", x)
  x <- trimws(x)
  gsub("\\s+", " ", x)
}

norm_text <- function(x) {
  tolower(clean_text(x))
}

clean_study_id <- function(x) {
  x <- clean_text(x)
  x <- gsub("\\.xlsm$", "", x, ignore.case = TRUE)
  x <- gsub("^v[0-9]+-", "", x, ignore.case = TRUE)
  x <- gsub("^v1002-", "", x, ignore.case = TRUE)
  x
}

is_real_study_id <- function(x) {
  grepl("^(BS|JM|JO|CJ)[0-9]+$", clean_study_id(x), ignore.case = TRUE)
}

get_prefix <- function(x) {
  toupper(sub("^([A-Za-z]+).*", "\\1", clean_study_id(x)))
}

make_expected_file <- function(x) {
  paste0("v1002-", clean_study_id(x), ".xlsm")
}

extract_study_id_from_file <- function(file_name) {
  x <- basename(file_name)
  x <- gsub("\\.xlsm$", "", x, ignore.case = TRUE)
  x <- gsub("^v[0-9]+-", "", x, ignore.case = TRUE)
  x
}

normalise_status <- function(x) {
  x <- norm_text(x)
  x[x %in% c("complete")] <- "completed"
  x[x %in% c("na", "n.a", "n.a.", "n/a")] <- "n/a"
  x[x %in% c("reject")] <- "rejected"
  x[x %in% c("queue")] <- "queued"
  x
}

list_xlsm_files <- function(folder) {
  if (!dir.exists(folder)) return(character(0))

  list.files(
    folder,
    pattern = "^v[0-9]+.*\\.xlsm$",
    full.names = TRUE,
    recursive = TRUE
  )
}

bucket_to_folder <- function(prefix, bucket) {
  subfolder <- if (bucket == "EXTRACTED") {
    "2.Extracted"
  } else if (bucket == "QC") {
    "3.Quality control"
  } else if (bucket == "QUEUED") {
    "4.Queued"
  } else if (bucket == "REJECTED") {
    "5.Rejected"
  } else {
    NA_character_
  }

  file.path(base_paths[[prefix]], subfolder)
}

get_expected_bucket <- function(in_qc_sheet, qc_status, comments_addressed, final_check) {
  qc_status <- normalise_status(qc_status)
  comments_addressed <- norm_text(comments_addressed)
  final_check <- norm_text(final_check)

  # Not yet listed in QC sheet
  if (!in_qc_sheet) return("EXTRACTED")

  # Direct QC outcomes
  if (qc_status == "rejected") return("REJECTED")
  if (qc_status == "queued") return("QUEUED")

  # Fully completed QC
  if (qc_status == "completed" && final_check == "yes-all") return("QC")

  # Completed QC with zero/no comments to address
  if (
    qc_status == "completed" &&
      comments_addressed %in% c("", "n/a", "na", "zero errors", "no comments")
  ) {
    return("QC")
  }

  # QC not yet complete, or comments still pending
  return("EXTRACTED")
}

# -----------------------------
# Match sheet names safely
# -----------------------------

if (!file.exists(progress_file)) {
  stop("Progress tracker not found. Check progress_file path.")
}

available_sheets <- excel_sheets(progress_file)
sheet_lookup <- setNames(available_sheets, trimws(available_sheets))

extract_sheets <- unname(sheet_lookup[extract_sheets_requested])
qc_sheets <- unname(sheet_lookup[qc_sheets_requested])

extract_sheets <- extract_sheets[!is.na(extract_sheets)]
qc_sheets <- qc_sheets[!is.na(qc_sheets)]

if (length(extract_sheets) == 0) {
  stop("No extract sheets found.")
}

# -----------------------------
# Read Extract sheets
# -----------------------------

extract_list <- list()

for (sheet in extract_sheets) {
  message("Reading extract sheet: ", sheet)

  dat <- suppressMessages(read_excel(progress_file, sheet = sheet, col_names = FALSE))
  if (nrow(dat) < 5) next

  # Row 1 = title
  # Row 2 = grouped headers
  # Row 3 = field names
  # Row 4 = guidance row
  # Row 5 onward = actual data
  dat <- dat[-c(1:4), ]

  tmp <- data.frame(
    extract_sheet = sheet,
    unique_study_id = clean_study_id(dat[[1]]),      # Column A
    new_tool_version = clean_text(dat[[4]]),         # Column D
    extraction_status = normalise_status(dat[[9]]),  # Column I
    stringsAsFactors = FALSE
  )

  tmp <- tmp[
    is_real_study_id(tmp$unique_study_id) &
      tmp$new_tool_version == "v1002" &
      tmp$extraction_status == "completed",
  ]

  if (nrow(tmp) > 0) {
    extract_list[[length(extract_list) + 1]] <- tmp
  }
}

if (length(extract_list) == 0) {
  stop("No completed v1002 extracted records found.")
}

extract_done <- unique(do.call(rbind, extract_list))

# -----------------------------
# Read QC sheets
# -----------------------------

qc_list <- list()

for (sheet in qc_sheets) {
  message("Reading QC sheet: ", sheet)

  dat <- suppressMessages(read_excel(progress_file, sheet = sheet, col_names = FALSE))
  if (nrow(dat) < 5) next

  dat <- dat[-c(1:4), ]

  qcer_from_sheet <- toupper(sub("-QC$", "", trimws(sheet)))

  tmp <- data.frame(
    qc_sheet = sheet,
    qcer_from_sheet = qcer_from_sheet,
    unique_study_id = clean_study_id(dat[[1]]),      # Column A
    extractor_recorded = clean_text(dat[[3]]),       # Column C
    qc_status = normalise_status(dat[[7]]),          # Column G
    comments_addressed = norm_text(dat[[8]]),        # Column H
    final_qc_check = norm_text(dat[[9]]),            # Column I
    stringsAsFactors = FALSE
  )

  tmp <- tmp[is_real_study_id(tmp$unique_study_id), ]

  if (nrow(tmp) > 0) {
    qc_list[[length(qc_list) + 1]] <- tmp
  }
}

qc_statuses <- if (length(qc_list) > 0) {
  unique(do.call(rbind, qc_list))
} else {
  data.frame(
    qc_sheet = character(),
    qcer_from_sheet = character(),
    unique_study_id = character(),
    extractor_recorded = character(),
    qc_status = character(),
    comments_addressed = character(),
    final_qc_check = character(),
    stringsAsFactors = FALSE
  )
}

# -----------------------------
# Merge extract-complete records with QC information
# -----------------------------

tracker <- merge(
  extract_done,
  qc_statuses,
  by = "unique_study_id",
  all.x = TRUE
)

tracker$in_qc_sheet <- !is.na(tracker$qc_sheet) & tracker$qc_sheet != ""

tracker$qc_status[is.na(tracker$qc_status)] <- ""
tracker$comments_addressed[is.na(tracker$comments_addressed)] <- ""
tracker$final_qc_check[is.na(tracker$final_qc_check)] <- ""
tracker$qcer_from_sheet[is.na(tracker$qcer_from_sheet)] <- ""

tracker$expected_prefix <- vapply(tracker$unique_study_id, get_prefix, character(1))
tracker$expected_file <- vapply(tracker$unique_study_id, make_expected_file, character(1))

tracker$expected_bucket <- mapply(
  get_expected_bucket,
  tracker$in_qc_sheet,
  tracker$qc_status,
  tracker$comments_addressed,
  tracker$final_qc_check
)

tracker$expected_folder <- mapply(
  bucket_to_folder,
  tracker$expected_prefix,
  tracker$expected_bucket
)

tracker$own_qc_flag <- ifelse(
  tracker$in_qc_sheet &
    tracker$qcer_from_sheet == tracker$expected_prefix,
  "POSSIBLE_SELF_QC",
  ""
)

# -----------------------------
# Build folder index
# -----------------------------

folder_index <- data.frame()

for (prefix in names(base_paths)) {
  folders_to_check <- data.frame(
    prefix = prefix,
    bucket = c("EXTRACTED", "QC", "QUEUED", "REJECTED"),
    folder = file.path(
      base_paths[[prefix]],
      c("2.Extracted", "3.Quality control", "4.Queued", "5.Rejected")
    ),
    stringsAsFactors = FALSE
  )

  for (j in seq_len(nrow(folders_to_check))) {
    files_found <- list_xlsm_files(folders_to_check$folder[j])

    if (length(files_found) > 0) {
      folder_index <- rbind(
        folder_index,
        data.frame(
          found_prefix = folders_to_check$prefix[j],
          found_bucket = folders_to_check$bucket[j],
          found_folder = folders_to_check$folder[j],
          found_file = basename(files_found),
          found_study_id = extract_study_id_from_file(files_found),
          found_path = files_found,
          stringsAsFactors = FALSE
        )
      )
    }
  }
}

# -----------------------------
# Compare expected folder with actual folder
# -----------------------------

inventory_list <- list()

for (i in seq_len(nrow(tracker))) {
  row <- tracker[i, ]

  # Match on paper ID, not file version
  found <- folder_index[
    tolower(folder_index$found_study_id) == tolower(row$unique_study_id),
  ]

  if (nrow(found) == 0) {

    placement_status <- "MISSING_FROM_ALL_TRACKED_FOLDERS"
    found_prefix <- found_bucket <- found_folder <- found_file <- found_path <- NA_character_

  } else if (any(
    found$found_prefix == row$expected_prefix &
      found$found_bucket == row$expected_bucket
  )) {

    placement_status <- "FOUND"

    good <- found[
      found$found_prefix == row$expected_prefix &
        found$found_bucket == row$expected_bucket,
    ][1, ]

    found_prefix <- good$found_prefix
    found_bucket <- good$found_bucket
    found_folder <- good$found_folder
    found_file <- good$found_file
    found_path <- good$found_path

  } else {

    placement_status <- "MISPLACED"
    found_prefix <- paste(unique(found$found_prefix), collapse = "; ")
    found_bucket <- paste(unique(found$found_bucket), collapse = "; ")
    found_folder <- paste(unique(found$found_folder), collapse = "; ")
    found_file <- paste(unique(found$found_file), collapse = "; ")
    found_path <- paste(unique(found$found_path), collapse = "; ")
  }

  inventory_list[[length(inventory_list) + 1]] <- data.frame(
    unique_study_id = row$unique_study_id,
    expected_file = row$expected_file,
    extract_sheet = row$extract_sheet,
    in_qc_sheet = row$in_qc_sheet,
    qc_sheet = row$qc_sheet,
    qcer_from_sheet = row$qcer_from_sheet,
    qc_status = row$qc_status,
    comments_addressed = row$comments_addressed,
    final_qc_check = row$final_qc_check,
    expected_prefix = row$expected_prefix,
    expected_bucket = row$expected_bucket,
    expected_folder = row$expected_folder,
    placement_status = placement_status,
    own_qc_flag = row$own_qc_flag,
    found_prefix = found_prefix,
    found_bucket = found_bucket,
    found_folder = found_folder,
    found_file = found_file,
    found_path = found_path,
    stringsAsFactors = FALSE
  )
}

inventory <- do.call(rbind, inventory_list)

issues <- inventory[
  inventory$placement_status %in% c("MISSING_FROM_ALL_TRACKED_FOLDERS", "MISPLACED") |
    inventory$own_qc_flag == "POSSIBLE_SELF_QC",
]

write.csv(issues, output_file, row.names = FALSE)

cat("\nDone.\n")
cat("Extracted-completed v1002 records checked:", nrow(tracker), "\n")
cat("Issues found:", nrow(issues), "\n")
cat("Output saved to:", output_file, "\n")

