---
title: "Duplicate XLSM File ID Check"
author: "Charity Juma"
date: "2026-05-18"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE
)
```

# Overview

This report scans multiple project folders for `.xlsm` files and identifies duplicate file IDs based on the last 5 digits found before the `.xlsm` extension.

The workflow:

1. Scan all specified folders recursively
2. Extract the final 5-digit identifier from filenames
3. Identify duplicate IDs
4. Export outputs to CSV files

# Load Folder Paths

```{r folders}
# Root folders to scan
folders <- c(
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Charity/2.Extracted",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Charity/3.Quality control",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Boris/2.Extracted",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Boris/3.Quality control",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Jabesh/2.Extracted",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Jabesh/3.Quality control",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Judy/2.Extracted",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Judy/3.Quality control",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Jabesh/1.Work in progress",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Judy/1.Work in progress",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Charity/1.Work in progress",
  "C:/Users/cjuma/CGIAR/Nowak, Andreea (Alliance Bioversity-CIAT) - AE-SysRev-shared/Boris/1.Work in progress"
)
```

# Collect XLSM Files

This section searches all folders recursively and lists all `.xlsm` files.

```{r collect-files}
files <- unlist(lapply(folders, function(r) {
  list.files(
    r,
    pattern = "\\.xlsm$",
    full.names = TRUE,
    recursive = TRUE
  )
}), use.names = FALSE)

cat("Total .xlsm files found:", length(files), "\n")
```

# Extract Last 5-Digit ID

The following function extracts the last 5 digits before the `.xlsm` extension.

```{r extract-ids}
get_last5 <- function(path) {
  
  base <- basename(path)
  
  m <- regexpr("(\\d{5})(?=\\.xlsm$)", base, perl = TRUE)
  
  if (m[1] == -1) {
    return(NA_character_)
  }
  
  regmatches(base, m)
}

ids5 <- vapply(files, get_last5, FUN.VALUE = character(1))
```

# Build File Table

Create a data frame containing:

- Extracted ID
- File name
- Full file path

```{r build-table}
df <- data.frame(
  id_last5 = ids5,
  file_name = basename(files),
  file_path = files,
  stringsAsFactors = FALSE
)

head(df)
```

# Files Without Detectable IDs

Display files that do not contain a valid 5-digit ID before `.xlsm`.

```{r missing-ids}
no_id <- df[is.na(df$id_last5), ]

if (nrow(no_id) > 0) {
  
  cat("Files with NO last-5-digit ID detected:\n")
  
  no_id[, c("file_name", "file_path")]
  
} else {
  
  cat("All files contain detectable IDs.")
}
```

# Detect Duplicate IDs

Identify duplicate 5-digit IDs appearing in multiple files.

```{r duplicates}
df_ok <- df[!is.na(df$id_last5), ]

dup_ids <- names(which(table(df_ok$id_last5) > 1))

dups <- df_ok[df_ok$id_last5 %in% dup_ids, ]

dups <- dups[order(dups$id_last5, dups$file_name), ]

cat("Duplicate ID groups found:", length(dup_ids), "\n")
```

# Duplicate File Results

```{r show-duplicates}
dups
```

# Export Results

Save outputs as CSV files.

```{r export-results}
write.csv(
  df_ok,
  "all_files_with_last5_id.csv",
  row.names = FALSE
)

write.csv(
  dups,
  "duplicate_last5_ids.csv",
  row.names = FALSE
)

cat("CSV exports completed successfully.")
```

# Summary

This report:

- Scanned all specified folders
- Extracted 5-digit identifiers from `.xlsm` files
- Detected duplicate IDs
- Exported results for review

Generated output files:

- `all_files_with_last5_id.csv`
- `duplicate_last5_ids.csv`
